Online Laterality Battery Pre-Processing - short version

### 2020-07-18 battery-preprocessing_v6_DB.Rmd
1. Rather than creating separate for_figs.csv file, add the L and R raw scores to allsum for all variables. 2. Remove all analyses for raw RT in RDT, since we preregistered use of log RT scores. Have done this now by just substituting logRT at an early stage and deleting separate section for logs. As this involves deleting chunks of stuff I have created new version of this script. 3. For DL2 there were 7 cases with very few trials - this looked like a technical error. I removed their data - see chunk where techerrs is defined. 4. The t-test laterality index for RDT had got lost somewhere in tr

@@

09 June 2020: AP made changes to calcultaions of rhyme
(1) Added t scores to output
(2) computed LIs for log-RT
(3) changed polariies so that a negative value was indicative of left lateralisation in the brain.

@@

# recode variabels
EHI$subject <- EHI$Participant.Public.ID
EHI$subject <- as.factor(EHI$subject)

@@

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "https://osf.io/mhk29/download" # day 1 data

  if (d ==2){
    data_file <-  "https://osf.io/ebm6s/download" # day 2 data
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

@@

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for sIDe
    temp_dat$accurate[myrow] <- sum(tmp1$Correct,na.rm=TRUE) # add row for N accurate response
    temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice. For RDT, can make error of selecting wrong sIDe, so % correct is meaningful

@@

# lets keep a running check on the participants completing here
complete.day2.CF <- unique(CF_dat2$subject)
```

@@

origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```

@@

# once again, mark rather than remove those with poor performance on identical stimuli
w<-union(which(allsum$sound.recog1 < .75), which(allsum$sound.recog2 < .75))
ww<-intersect(which(allsum$exclude==1),w)
#ww contains people who were excluded on both dichotic and faces. 
# code exclude = 2 for the cases where there less than 75% for identical stimuli
allsum$exclude[w]<-2 #code exclude = 2 if excluded because poor dichotic
allsum$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
t <- table(allsum$exclude) 
rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich')
t
```

@@

```{r dosummarydich}
includesubs <- which(allsum$exclude<2) #only plot/t-test included cases
origdat <- DL_dat1[DL_dat1$subject %in% includesubs,]
measure <- 'Accuracy'
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 1'
dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
t.test(origdat$accurate~ origdat$side)
#Now repeat for day 2
origdat <-  DL_dat2[DL_dat2$subject %in% includesubs,]
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```

@@

allsum$exclude[w]<-2 #mark as excluded for dichotic

@@

```{r rdt_outliers}
RDT_dat1<-detect.outliers(RDT_dat1,200,1.65) #numbers specify min RT and 
RDT_dat2<-detect.outliers(RDT_dat2,200,1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively

@@

```{r makedf_rhyme}
#We create chim_dat files for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
latlist <- c("left","no rhyme","right") #names of factor levels
nloop<-length(latlist)
RDT_df1 <- make.df(RDT_dat1,varlist,latlist,nloop)
RDT_df2 <- make.df(RDT_dat2,varlist,latlist,nloop)

@@

Comparison shows that there is significant difference in RT. Accuracy is close to ceiling and does not show side difference (but it would be a bit odd for this task, - unlike CF and dichotic, there is no competing information and only one answer is correct).
```{r rdtpirate}
for (d in 1:2){
origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
origdat <-origdat[origdat$side %in% c('left','right'),]

@@

@@

# RT
origdat$X1 <- origdat$RT
mycolname<-"Day1.RDT_RT_LI"
if(d==2){mycolname<-"Day2.RDT_RT_LI"}
polarity <- (1)
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
}

@@

# correlate
cor1<-cor(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,use='complete.obs')
# acc plot
plot(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
# raw plot
plot(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,9,paste0('r = ',round(cor2,3)))
```
## Does RDT relate to handedness?
Answer is convincing nope!

@@

### T-test as alternative to LI for RDT

The LI results are not encouraging - reliability well below acceptable level, and although RT shows lateralised bias, it is tiny. Also, many subjects dropped.
Previously I also did a t-test that was analogous to the z-test for CF/dichotic, i.e. showed if the individual had a significant RT difference between L and R stimuli. But that is v highly correlated with LI. The only way to make it more sensitive would be to try a pairwise t-test that compares RTs for the same items presented on L and R. So the next bit of code is designed to compute that.

```{r RDTaggregate}
#remove 'no rhyme' trials
for (d in 1:2){
  thisdat<-RDT_dat2[RDT_dat2$side %in% c('left','right'),]
  if(d==1){
    thisdat<-RDT_dat1[RDT_dat1$side %in% c('left','right'),]
  }

  thisdat <- thisdat[thisdat$outlier==0,] #remove outliers
  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$Participant.Public.ID),
                       FUN=mean,na.rm=TRUE)
  aggdata$subject <- aggdata$Participant.Public.ID
  colnames(aggdata)<-c('side','item','subject','RT')
  mysubs<-unique(aggdata$subject)
  if(d==1){
  pairedt <- data.frame(matrix(NA,nrow=length(mysubs),ncol=5))
  colnames(pairedt) <- c('subject','t1','p1','t2','p2')
  }
  nsubs<-length(mysubs)
  for (i in 1:nsubs){
    myt<-NA
    myp<-NA
    thisdat <- aggdata[aggdata$subject==mysubs[i],]
    pairedt$subject[i]<- mysubs[i]
    #check there are data for both left and right on all items
    checkdat <-table(thisdat$item)
    w<-which(checkdat<2)
    if(length(w)==0){
      t<-t.test(thisdat$RT[thisdat$side=='left'],thisdat$RT[thisdat$side=='right'],paired=T)
      myt<-t$statistic
      myp<-t$p.value
    }
    if(d==1){
      pairedt$t1[i] <- myt
      pairedt$p1[i] <-myp
    }
    if(d==2){
      pairedt$t2[i] <-myt
      pairedt$p2[i] <-myp
    }
  }
}

# add t score to the data
t_dat <- select(pairedt, -3, -5)
t_dat <- na.omit(t_dat)
# merge to allsum for analysis
allsum <- merge(allsum, t_dat, by.x = "origID", by.y = "subject", all.x=T)

cor(pairedt$t1,pairedt$t2,use='complete.obs')
```

So the paired items analysis gives a totally nonsignificant correlation between time 1 and time 2. This is much worse than the overall LI reliability.

The failure of this task led to further thoughts about why it is barely lateralised, given that it does implicate language-specific word generation. 
A noteworthy feature of the other two tasks (CF and dichotic) is that they measure side preference rather than accuracy - i.e. a response from either L or R is correct, but two potential responses are put in competition. That contrasts with this task where only one side has the correct answer.  
We used pictures because with written words there is a concern that  preference may be confounded by the fact we read words from L to R. That could potentially dilute the RVF advantage that is expected from L hem language.

### log RT

In our pre-reg, we also specified that we would use logRT. So we add this in here.

```{r logRT analysis}
# now add the logRT to original output
# first relabel df
RDT_logdat1 <- RDT_dat1
RDT_logdat2 <- RDT_dat2
# now log the RT
RDT_logdat1$RT <- log(RDT_dat1$RT)
RDT_logdat2$RT <- log(RDT_dat2$RT)
# Now use generic function to remove outliers.
# Here we keep the 1.65 of the H&I removal. 
# BUT we now specify log 200 for lower cutoff to move onto the same scale
RDT_logdat1<-detect.outliers(RDT_logdat1,log(200),1.65) #numbers specify min RT and 
RDT_logdat2<-detect.outliers(RDT_logdat2,log(200),1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectiv
# We create logRT for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
latlist <- c("left","no rhyme","right") #names of factor levels
nloop<-length(latlist)
RDT_log_df1 <- make.df(RDT_logdat1,varlist,latlist,nloop)
RDT_log_df2 <- make.df(RDT_logdat2,varlist,latlist,nloop)
# remove NaN
is.nan.data.frame <- function(x)
  do.call(cbind, lapply(x, is.nan))
RDT_log_df1[is.nan(RDT_log_df1)] <- NA
RDT_log_df2[is.nan(RDT_log_df2)] <- NA
# now let's plot the log data
for (d in 1:2){
origdat <-RDT_log_df1[RDT_log_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_log_df2[RDT_log_df2$subject %in% incRDT,]}
origdat <-origdat[origdat$side %in% c('left','right'),]
# RT
origdat$X1 <- origdat$RT
measure <- 'log-RT'
dopirate(origdat,task,measure)
}
# make LI for log RT
# Because outliers will already be marked in allsum, we won't bother going through this again. 
# We will intead add straight into the data frame. 
#now create LI
for (d in 1:2){
origdat <-RDT_log_df1[RDT_log_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_log_df2[RDT_log_df2$subject %in% incRDT,]}
# log RT
origdat$side <- as.factor(origdat$side)
  levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 0
origdat$X1 <- origdat$RT
mycolname<-"Day1_RDT_logRT_LI"
if(d==2){mycolname<-"Day2_RDT_logRT_LI"}
polarity <- (1)
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
}
# plot log RT
cor1<-cor(allsum$Day1_RDT_logRT_LI,allsum$Day2_RDT_logRT_LI,use='complete.obs')
plot(allsum$Day1_RDT_logRT_LI,allsum$Day2_RDT_logRT_LI,main='Rhyme detect: LI for log RT')
abline(v=0)
abline(h=0)
text(-1,1,paste0('r = ',round(cor1,3)))
# FINALLY, make t-test for logRT
#remove 'no rhyme' trials
for (d in 1:2){
  thisdat<-RDT_logdat2[RDT_logdat2$side %in% c('left','right'),]
  if(d==1){
    thisdat<-RDT_logdat1[RDT_logdat1$side %in% c('left','right'),]
  }

  thisdat <- thisdat[thisdat$outlier==0,] #remove outliers
  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$Participant.Public.ID),
                       FUN=mean,na.rm=TRUE)
  aggdata$subject <- aggdata$Participant.Public.ID
  colnames(aggdata)<-c('side','item','subject','RT')
  mysubs<-unique(aggdata$subject)
  if(d==1){
  pairedt <- data.frame(matrix(NA,nrow=length(mysubs),ncol=5))
  colnames(pairedt) <- c('subject','t1','p1','t2','p2')
  }
  nsubs<-length(mysubs)
  for (i in 1:nsubs){
    myt<-NA
    myp<-NA
    thisdat <- aggdata[aggdata$subject==mysubs[i],]
    pairedt$subject[i]<- mysubs[i]
    #check there are data for both left and right on all items
    checkdat <-table(thisdat$item)
    w<-which(checkdat<2)
    if(length(w)==0){
      t<-t.test(thisdat$RT[thisdat$side=='left'],thisdat$RT[thisdat$side=='right'],paired=T)
      myt<-t$statistic
      myp<-t$p.value
    }
    if(d==1){
      pairedt$t1[i] <- myt
      pairedt$p1[i] <-myp
    }
    if(d==2){
      pairedt$t2[i] <-myt
      pairedt$p2[i] <-myp
    }
  }
}

# add t score to the data
t_dat <- select(pairedt, -3, -5)
t_dat <- na.omit(t_dat)
# relabel for merging
names(t_dat)[names(t_dat) == "t1"] <- "logt1"
names(t_dat)[names(t_dat) == "t2"] <- "logt2"
# merge to allsum for analysis
allsum <- merge(allsum, t_dat, by.x = "origID", by.y = "subject", all.x=T)

cor(pairedt$t1,pairedt$t2,use='complete.obs')
```

Even when log-transforming, things don't look any better.

@@

In the raw data file, farticipants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

@@

    wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
    for (s in 1:2){
      myc<-mycols[d,s] #select column
      target <- myseq[s] #tumb then wrvx, ie right then left
      finger.df[n,myc]<-str_count(wholeseq,target) 
      #count how many target sequences are in the long string that contains all responses

    }
    #We then add the SD of the RT - this indicates variability in pace of keypressing
    righth<-filter(tempfinger,display=='right hand')
    lefth<-filter(tempfinger,display=='left hand')
    myc<-mycols[d,1]+4

    finger.df[n,myc]<-sd(righth$RTa,na.rm=T)
    finger.df[n,(myc+2)]<-sd(lefth$RTa,na.rm=T)
  }
}
# Substitute NA for any totals less than 5 - suggest using wrong keys perhaps

@@

origdat <- finger.long
measure <- '2sd'
origdat$X1 <- origdat$`2sd` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 2sd'
dopirate(origdat,task,measure)
```

@@

```{r fingerLI}
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_finger_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

cor1 <- cor(allsum$Day1_finger_LI,allsum$Day2_finger_LI,use='complete.obs')
plot(allsum$Day1_finger_LI,allsum$Day2_finger_LI,main='Finger Tapping LI')
abline(v=0)
abline(h=0)
text(-20,10,paste0('r = ',round(cor1,3)))
```

@@

task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_Finger_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-finger.long
origdat$X1 <- origdat$`2`
mycolname<-"Day2_Finger_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,use='complete.obs')
plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy',col=allsum$handpch,pch=15)
abline(v=0)
abline(h=0)
text(-2.5,4,paste0('r = ',round(cor1,3)))

@@

# now calculate % correct
plot.data <- plot.data %>%
  mutate(remove= if_else(plot.data$task== "RD" & plot.data$p.corr < .75,1,
                         if_else(plot.data$accurate <= 10,1,0)))
plot.data <- plot.data[plot.data$remove== 0,]

# save data frame
write.csv(plot.data, "./data/for_figs.csv")
```

@@

```{r reducerDL}
for (d in 1:2){ #we will process day 1 and day 2 together
  dat <- dichotic_dat1
  if (d==2)
  {dat <- dichotic_dat2}

>>>>

18th July DB edits
1. Rather than creating separate for_figs.csv file, add the L and R raw scores to allsum for all variables.
2. Remove all analyses for raw RT in RDT, since we preregistered use of log RT scores. Have done this now by just substituting logRT at an early stage and deleting separate section for logs. As this involves deleting chunks of stuff I have created new version of this script.
3. For DL2 there were 7 cases with very few trials - this looked like a technical error. I removed their data - see chunk where techerrs is defined.
4. The t-test laterality index for RDT had got lost somewhere in translation. I've put it back. It actually gives a test-retest reliability of .63! Ho hum.

@@

09 June 2020: AP made changes to calculations of rhyme
(1) Added t scores to output
(2) computed LIs for log-RT
(3) changed polariies so that a negative value was indicative of left lateralisation in the brain.

@@

# recode variables
EHI$subject <- EHI$Participant.Public.ID
EHI$subject <- as.factor(EHI$subject)

@@

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "https://osf.io/mhk29/download" #day 1 data

  if (d ==2){
    data_file <-  "https://osf.io/ebm6s/download"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

@@

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for sIDe
    temp_dat$accurate[myrow] <- sum(tmp1$Correct,na.rm=TRUE) # add row for N accurate response
    temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice. For RDT, can make error of selecting wrong side, so % correct is meaningful

@@

# lets keep a running check on the participants completing here

complete.day2.CF <- unique(CF_dat2$subject)

@@

origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)

#added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
chimdatL1<-select(filter(chim_dat1,side=='Left'),subject,accurate)
chimdatR1<-select(filter(chim_dat1,side=='Right'),subject,accurate)
chimdatL2<-select(filter(chim_dat2,side=='Left'),subject,accurate)
chimdatR2<-select(filter(chim_dat2,side=='Right'),subject,accurate)

allsum <- merge(allsum,chimdatL1,by="subject")
allsum <- merge(allsum,chimdatR1,by="subject")
allsum <- merge(allsum,chimdatL2,by="subject")
allsum <- merge(allsum,chimdatR2,by="subject")
myn <-ncol(allsum)
colnames(allsum)[(myn-3):myn] <- c('CF1.Lraw','CF1.Rraw','CF2.Lraw','CF2.Rraw')

@@

# once again, mark rather than remove those with poor performance on identical stimuli
w<-union(which(allsum$sound.recog1 < .75), which(allsum$sound.recog2 < .75))

ww<-intersect(which(allsum$exclude==1),w)
#Also exclude if less than 75% accuracy on the different trials

#ww contains people who were excluded on both dichotic and faces. 
# code exclude = 2 for the cases where there less than 75% for identical stimuli
allsum$exclude[w]<-2 #code exclude = 2 if excluded because poor dichotic
allsum$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
t <- table(allsum$exclude) 
rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich')
t
```

@@

We have to remove cases where v few trials because of technical problems. Specify minimum of 80 (out of 120) trials. (Most detected by this proces have less than 10 trials!).
NB This is only run for session 2 because there were no technical problems for session 1.

```{r marktecherr}
techerr<-table(dichotic_dat2$subject)
te<-which(techerr<80)
techerr<-techerr[te] #subject numbers for those with few trials.
techerr<-as.numeric(names(techerr))
```
The data is visualised using the pirate plot function and t.test conducted.

```{r dosummarydich}
includesubs <- which(allsum$exclude<2) #only plot/t-test included cases
origdat <- DL_dat1[DL_dat1$subject %in% includesubs,]
measure <- 'Accuracy'
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 1'
dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
t.test(origdat$accurate~ origdat$side)
#Now repeat for day 2
origdat <-  DL_dat2[DL_dat2$subject %in% includesubs,]
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```
```{r add.dichotic.raw}
#added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
DLdatL1<-select(filter(DL_dat1,side=='Left'),subject,accurate)
DLdatR1<-select(filter(DL_dat1,side=='Right'),subject,accurate)
DLdatL2<-select(filter(DL_dat2,side=='Left'),subject,accurate)
DLdatR2<-select(filter(DL_dat2,side=='Right'),subject,accurate)

#Add to allsum
allsum <- merge(allsum,DLdatL1,by="subject")
allsum <- merge(allsum,DLdatR1,by="subject")
allsum <- merge(allsum,DLdatL2,by="subject")
allsum <- merge(allsum,DLdatR2,by="subject")
myn <-ncol(allsum)
colnames(allsum)[(myn-3):myn] <- c('DL1.Lraw','DL1.Rraw','DL2.Lraw','DL2.Rraw')
#remove cases with technical errors from session 2
allsum$DL2.Lraw[techerr]<-NA
allsum$DL2.Rraw[techerr]<-NA

@@

allsum$exclude[w]<-2 #mark as excluded for dichotic
#Now remove values for the cases with technical error (v few trials)
allsum$Day2_DL_RT_LI[techerr] <-NA
allsum$Day2_DL_acc_LI[techerr] <-NA

@@

```{r rdt_outliers}
#First we create a RT column that is based on LOGRT
RDT_dat1$rawRT <- RDT_dat1$RT
RDT_dat2$rawRT <- RDT_dat2$RT
RDT_dat1$RT <- log(RDT_dat1$rawRT)
RDT_dat2$RT <- log(RDT_dat2$rawRT)
RDT_dat1<-detect.outliers(RDT_dat1,log(200),1.65) #numbers specify min RT and 
RDT_dat2<-detect.outliers(RDT_dat2,log(200),1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively

@@

```{r makedf_rhyme}
#We create RDT files for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
latlist <- c("left","no rhyme","right") #names of factor levels
nloop<-length(latlist)
RDT_df1 <- make.df(RDT_dat1,varlist,latlist,nloop)
RDT_df2 <- make.df(RDT_dat2,varlist,latlist,nloop)

@@

Comparison shows that there is significant side difference in RT. Accuracy is close to ceiling and does not show side difference (but it would be a bit odd for this task, - unlike CF and dichotic, there is no competing information and only one answer is correct).
```{r rdtpirate}
for (d in 1:2){
origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
origdat <-origdat[origdat$side %in% c('left','right'),]

@@

```{r add.rdt.raw}
#added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
RDTdatL1<-select(filter(RDT_df1,side=='left'),subject,RT)
RDTdatR1<-select(filter(RDT_df1,side=='right'),subject,RT)
RDTdatL2<-select(filter(RDT_df2,side=='left'),subject,RT)
RDTdatR2<-select(filter(RDT_df2,side=='right'),subject,RT)

allsum <- merge(allsum,RDTdatL1,by="subject")
allsum <- merge(allsum,RDTdatR1,by="subject")
allsum <- merge(allsum,RDTdatL2,by="subject")
allsum <- merge(allsum,RDTdatR2,by="subject")
myn <-ncol(allsum)
colnames(allsum)[(myn-3):myn] <- c('RDTlog1.Lraw','RDTlog1.Rraw','RDTlog2.Lraw','RDTlog2.Rraw')

@@

# RT
origdat$X1 <- origdat$RT
mycolname<-"Day1.RDT_RT_LI"
if(d==2){mycolname<-"Day2.RDT_RT_LI"}
polarity <- (1)
mycutoff <- 0 #no cutoff for logRT data
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
}

@@

# correlate
cor1<-cor(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,use='complete.obs')
# acc plot
plot(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
# raw plot
plot(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-1,1,paste0('r = ',round(cor2,3)))
```
### T-test as alternative to LI for RDT

Advantage of using a t-statistics to represent LI is that we can also identify people whose L-R difference is significantly different from zero.

```{r do.ttest}
for (d in 1:2){
  myfile<-RDT_dat1
  if(d==2){
    myfile<-RDT_dat2}

  myfile<-filter(myfile,side %in% c('left','right'))

  nsubs <- nrow(allsum)
  allsum$tlat_RDT<-NA

  for (i in 1:nsubs) {
    w<-which(myfile$subject==allsum$subject[i])
    if(length(w)>10)
    {
      myt<-t.test(myfile$RT[w]~myfile$side[w]) #this is based on logRT
      allsum$tlat_RDT[i]<-myt$statistic
    }
  }
  if (d==1){
    maxcol <-length(colnames(allsum))
    colnames(allsum)[maxcol]<-c('tlat_RDT_1')
  }
  if (d==2){
    maxcol <-length(colnames(allsum))
    colnames(allsum)[maxcol]<-c('tlat_RDT_2')
  }
}

colcount<-length(colnames(allsum))

w<-which(is.na(allsum$Day1.RDT_RT_LI)) #remove those with NA in LI
allsum[w,(colcount-1):colcount]<-NA
w<-which(is.na(allsum$Day2.RDT_RT_LI))
allsum[w,(colcount-1):colcount]<-NA

#compare t-value from day 1 and day 2 
myr <- cor(allsum$tlat_RDT_1,allsum$tlat_RDT_2,use='complete.obs')
plot(allsum$tlat_RDT_1,allsum$tlat_RDT_2,xlim=c(-5,5),ylim=c(-5,5))
abline(v=0)
abline(h=0)
abline(v=-1.65,lty=2,col='red') 
abline(h=-1.65,lty=2,col='red')
abline(v=1.65,lty=2,col='red') 
abline(h=1.65,lty=2,col='red')
text(-3,3,paste0('r = ',round(myr,3)))

plot(allsum$Day1.RDT_RT_LI,allsum$tlat_RDT_1,xlim=c(-1.5,1.5),ylim=c(-5,5))
```

## Does RDT relate to handedness?
Answer is convincing nope!

@@

@@

In the raw data file, participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

@@

    wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
    for (s in 1:2){ #cycle through sequences for L and R
      myc<-mycols[d,s] #select column
      target <- myseq[s] #tumb then wrvx, ie right then left
      mystrs<-str_count(wholeseq,target) 
      finger.df[n,myc]<-mystrs
      #count how many target sequences are in the long string that contains all responses
      if(s==1){mystrs0<-mystrs}
    }
    allseqd <-mystrs0+mystrs
    findbad <- abs(nchar(wholeseq)/4-allseqd) #allseqd x 4 should equal length of wholeseq if there are no wrong keypresses.
      if (findbad>20) #if more than 20 wrong keypresses, then exclude.
        {finger.df[n,mycols[d,1]]<-NA
        finger.df[n,mycols[d,2]]<-NA}
    #We then add the SD of the RT - this indicates variability in pace of keypressing
    righth<-filter(tempfinger,display=='right hand')
    lefth<-filter(tempfinger,display=='left hand')
    myc<-mycols[d,1]+4

    finger.df[n,myc]<-sd(righth$RTa,na.rm=T)
    finger.df[n,(myc+2)]<-sd(lefth$RTa,na.rm=T)
  }
}
# Substitute NA for any totals less than 5

@@

origdat <- finger.long
measure <- '2sd'
origdat$X1 <- origdat$`2sd` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 2sd'
dopirate(origdat,task,measure)
```
```{r add.fingerL_R}
#added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
fingerL1<-select(finger.df,subject,L1)
fingerR1<-select(finger.df,subject,R1)
fingerL2<-select(finger.df,subject,L2)
fingerR2<-select(finger.df,subject,R2)

allsum <- merge(allsum,fingerL1,by="subject")
allsum <- merge(allsum,fingerR1,by="subject")
allsum <- merge(allsum,fingerL2,by="subject")
allsum <- merge(allsum,fingerR2,by="subject")
myn <-ncol(allsum)
colnames(allsum)[(myn-3):myn] <- c('finger.L1','finger.R1','finger.L2','finger.R2')

@@

```{r fingerLI}
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_finger_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

cor1 <- cor(allsum$Day1_finger_LI,allsum$Day2_finger_LI,use='complete.obs')
plot(allsum$Day1_finger_LI,allsum$Day2_finger_LI,main='Finger Tapping LI',col=allsum$handpch,pch=15)
abline(v=0)
abline(h=0)
abline(0,1,lty=2)
text(-10,10,paste0('r = ',round(cor1,3)))
```

@@

task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_Finger_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-finger.long
origdat$X1 <- origdat$`2`
mycolname<-"Day2_Finger_acc_Z"
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,use='complete.obs')
plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy',col=allsum$handpch,pch=15)
abline(v=0)
abline(h=0)
abline(0,1,lty=2)
text(-2,3.5,paste0('r = ',round(cor1,3)))

@@

# save data frame
write.csv(plot.data, "./data/for_figs.csv")

@@

```{r reducerDL}
{for (d in 1:2){ #we will process day 1 and day 2 together
  dat <- dichotic_dat1
  if (d==2)
  {dat <- dichotic_dat2}
